---
title: "Using Linear Regression to Explain Deputies Votes"
author: "Jose Manoel Ferreira"
date: "1 de outubro de 2018"
output: html_document
---

``` {r prepare, cache=TRUE, message=FALSE, warning=FALSE}

  # campos:
  # "sequencial_candidato" :(character) id do candidato
  # "nome": (character)
  # "uf": (character)
  # "partido": (character)
  # "quantidade_doacoes": (integer)
  # "quantidade_doadores": (integer) número de doadores diferentes
  # "total_receita": (double) soma em R$ das doações
  # "media_receita": (double) média das doações
  # "recursos_de_outros_candidatos/comites": (double) quantia em R$ das doações provenientes de outros candidatos ou comite partidário
  # "recursos_de_pessoas_fisicas": (double) quantia em R$ das doações provenientes de outros CPFs
  # "recursos_de_pessoas_juridicas": (double) quantia em R$ das doações provenientes de outros CNPJ
  # "recursos_proprios": (double) quantia em R$ das doações provenientes do próprio candidato
  # "recursos_de_partido_politico": (double) quantia em R$ das doações provenientes do partido político do candidato
  # "votos": (integer) variável alvo. Se refere ao número de votos na campanha de 2006 e 2010
  # "quantidade_despesas": (integer)
  # "quantidade_fornecedores": (integer) número de fornecedores/despesas diferentes
  # "total_despesa": (double) soma em R$ das despesas de campanha
  # "media_despesa": (double) média das despesas de campanha
  # "cargo": (character)
  # "Sexo":  (character)
  # "grau": (character) grau de instrução do candidato
  # "estado_civil": (character)
  # "ocupacao": (character) ocupação do candidato
```

Setting up a workspace
```{r ,warning=FALSE, message=FALSE}
setwd("~/git/data-analysis/lab02/")
```

Building our dataframes

```{r ,warning=FALSE, message=FALSE}
data2006 <- read.csv(("data/eleicoes_2006.csv"))
data2010 <- read.csv(("data/eleicoes_2010.csv"))
data2006_2010 <-read.csv(("data/eleicoes_2006_e_2010.csv"))

```

Removing the NA

```{r ,warning=FALSE, message=FALSE}
data2006 <- na.omit(data2006)
data2010 <- na.omit(data2010)
data2006_2010 <- na.omit(data2006_2010)
```

Library used

```{r ,warning=FALSE, message=FALSE}
library(reshape)
library(ggplot2)
library(dplyr)
library(lattice)
library(scales)
library(corrplot)
library("PerformanceAnalytics")
library(GGally)
```

Converting some categoric variables to factor, basically it set up a specific number to each value.

#### 2006
```{r ,warning=FALSE, message=FALSE}
data2006$sexo <- as.factor(data2006$sexo)
data2006$estado_civil <- as.factor(data2006$estado_civil)
data2006$uf <- as.factor(data2006$uf)
data2006$sexo <- as.factor(data2006$sexo)
data2006$grau <- as.factor(data2006$grau)
data2006$estado_civil <- as.factor(data2006$estado_civil)
data2006$ocupacao <- as.factor(data2006$ocupacao)

```


#### 2010
```{r ,warning=FALSE, message=FALSE}
data2010$sexo <- as.factor(data2010$sexo)
data2010$estado_civil <- as.factor(data2010$estado_civil)
data2010$uf <- as.factor(data2010$uf)
data2010$sexo <- as.factor(data2010$sexo)
data2010$grau <- as.factor(data2010$grau)
data2010$estado_civil <- as.factor(data2010$estado_civil)
data2010$ocupacao <- as.factor(data2010$ocupacao)

```


Bar charts to numeric variables

#### 2006

```{r ,warning=FALSE, message=FALSE}

d = melt(select(data2006, -sequencial_candidato, - ano))
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable, scales = "free_x") + 
    scale_x_continuous(labels = comma) +
    geom_histogram()

```

#### 2010
```{r ,warning=FALSE, message=FALSE}
d = melt(select(data2010, -sequencial_candidato, - ano))
ggplot(d,aes(x = value)) + 
    facet_wrap(~variable, scales = "free_x") + 
    scale_x_continuous(labels = comma) +
    geom_histogram()
```


It's important to see the correlation between the variables, it will show to us the dependence between variables

#### 2006


```{r ,warning=FALSE, message=FALSE, fig.height=15,fig.width=15}
data2006.correlation1 <- data2006 %>% select(-c(sequencial_candidato, nome, estado_civil, ano, cargo))

data2006.correlation <- data2006.correlation1  %>%
  mutate(uf = as.numeric(uf)) %>%
  mutate(partido = as.numeric(partido)) %>%
  mutate(sexo = as.numeric(sexo)) %>%
  mutate(grau = as.numeric(grau)) %>%
  mutate(ocupacao = as.numeric(ocupacao))
  

ggpairs(data2006.correlation, axisLabels = "none")
```

Seeing the correlation betweend the variables we can say that the variable whit the best correlation whit votos is total_despesa 0.658 flollowed by total_receita 0.635 whitch makes 
totally sense, because the total_receita is the values used to pay the expense and its represented by total_despesas.
>
  * Greater correation is whit:
    > total_despesas corr: 0.658
    > total_receita corr: 0.635
> 
  * Smaller correlation is whit:
    > uf corr: - 0.0318
    > partido: - 0.048



#### 2010
votos is our target variable, so let's take a look and the correlation between votos and other variables.
>
  * Greater correation is whit:
  > total_despesas corr: 0.637
  > total_receita  corr: 0.626
> 
  * Smaller correlation is whit:
    > media_despesas corr:   0.0715
    > uf             corr: - 0.0425
  
  
  
#### 2006 and 2010
As we can see we've got a very similar result in both years, total_receita and total_despesa as the greater correlation and uf was the smalest.

```{r ,warning=FALSE, message=FALSE, fig.height=15,fig.width=15}
data2010.correlation1 <- data2010 %>% select(-c(sequencial_candidato, nome, estado_civil, ano, cargo))

data2010.correlation <- data2010.correlation1  %>%
  mutate(uf = as.numeric(uf)) %>%
  mutate(partido = as.numeric(partido)) %>%
  mutate(sexo = as.numeric(sexo)) %>%
  mutate(grau = as.numeric(grau)) %>%
  mutate(ocupacao = as.numeric(ocupacao))
  

ggpairs(data2010.correlation, axisLabels = "none")
```

Multiple Regression Linear Model Whit all numeric variables

### 1. A multiple regresson model whit all variables is able to explain the y variation (total of votos) in 2006 and 2010?

#### 2006
> As we can see we've got a $R_a^2$ and $R^2$ whit 0.5779 and  0.5488 respectively. As close to 1 is better to explain the model, but it seems not enough and even adjusting the R-squared we got a minimum increasing. So the model whit all variables does not weel explains the total of votos. Also there is too many categoric variables making an extense model being difficult to visualise.


```{r ,warning=FALSE, message=FALSE}

data2006.lm = lm(data2006$votos ~ data2006$uf + data2006$partido  + data2006$quantidade_doacoes + data2006$quantidade_doadores + data2006$total_receita + data2006$media_receita + data2006$recursos_de_outros_candidatos.comites + data2006$recursos_de_pessoas_fisicas + data2006$recursos_de_pessoas_juridicas + data2006$recursos_proprios + data2006$recursos_de_partido_politico + data2006$quantidade_despesas + data2006$quantidade_fornecedores + data2006$total_despesa + data2006$media_despesa + data2006$sexo + data2006$grau + data2006$estado_civil + data2006$ocupacao)

summary(data2006.lm)

```

#### 2010

Not different of 2006 the model whit all variables is not good to explaing the variable votos R-square was 0.5405 and	Adjusted R-squared:  0.5129. And it is interesting to point out that adjusting the R-square it decrease the value.

```{r ,warning=FALSE, message=FALSE}

data2010.lm = lm(data2010$votos ~ data2010$uf + data2010$partido  + data2010$quantidade_doacoes + data2010$quantidade_doadores + data2010$total_receita + data2010$media_receita + data2010$recursos_de_outros_candidatos.comites + data2010$recursos_de_pessoas_fisicas + data2010$recursos_de_pessoas_juridicas + data2010$recursos_proprios + data2010$recursos_de_partido_politico + data2010$quantidade_despesas + data2010$quantidade_fornecedores + data2010$total_despesa + data2010$media_despesa + data2010$sexo + data2010$grau + data2010$estado_civil + data2010$ocupacao)

summary(data2010.lm)

```

### 2. Compere the regressiond made to 2006 and 2010. Which are the difference/similarities noticed:

> Looking to the residual charts we can notice that the greater amount is close to zero, it is good because there is not big difference between the new model results copared whit the original one. But the scale to residual chart of 2010 is greter then 2006 it means that had some errors whit a considerable value.
> The variables has almost the same correlation value, it looks so same to both years.

  

#### 2006
```{r}
data2006.residuals = resid(data2006.lm) 

densityplot(data2006.residuals)
```


#### 2010
```{r}
data2010.residuals = resid(data2010.lm) 

densityplot(data2010.residuals)
```
### 3. Are all variables useful to the regression model? is there any redundant varialbe? Alaysis to 2006 and 2010 separately.

> No, not all variable are usefull to the model, for example cargo, there is only one cargo and year which is DEPUTADO FEDERAL and 2006 respectively. Also there is redudant variables, media_receita which is the mediam of all receita so it have a strong correlation whit total_receita, so it not necessary to have both variables in the model. Also we've some variables whit a strong correlation as quantidade_doacoes and quantidade_doadores whit 0.997 in both years; total_receita and total_despesas whit 0.99; total_receita and total_despesas whit  0.988, those variable together are not good to explain the model, so should chouse one of each variable strong correlated.



### 4. If there is redudant variables or not explained, make a new model excluding those variables and compare the new model with the one that has all variables (e.g. should explain using R2 and RSE)


> In order to reach a better model we've choose a few variables, numeric ones whit a good significance value, but the R-squared and adjusted got was smaller then the one whit all varialbes, also not a good regression model, not 2006 nor 2010.
  *  2006: R-squared:  0.4896,	Adjusted R-squared:  0.4887 
  *  2010: R-squared:  0.4232,	Adjusted R-squared:  0.4224 

#### 2006
```{r}
data2006.lm2 = lm(data2006$votos ~ data2006$quantidade_doadores + data2006$total_receita + data2006$recursos_de_pessoas_fisicas + data2006$recursos_de_pessoas_juridicas + data2006$quantidade_despesas + data2006$total_despesa)

summary(data2006.lm2)
```
#### 2010
```{r}
data2010.lm2 = lm(data2010$votos ~ data2010$quantidade_doadores + data2010$total_receita + data2010$recursos_de_pessoas_fisicas + data2010$recursos_de_pessoas_juridicas + data2010$quantidade_despesas + data2010$total_despesa)

summary(data2010.lm2)
```

### 5. Make a regression including both years 2006 and 1010 together. Which differencies/similitaries are notices coparing whit the models whit individual year? 
> Putting all year together and we notice that correlation does not change too much, and the year in all variables show a small correlation factor.
> Almost same residual chart, but more close to 2010.
> R-squared:  0.5082,	Adjusted R-squared:  0.4909 Was a few better then other year
> Even reproducing the model whit fewer variables, removing the redundant ones whe still getting a R2 and RSE between 0.4 and 0.5


```{r ,warning=FALSE, message=FALSE, fig.height=15,fig.width=15}
data2006_2010.correlation1 <- data2006_2010 %>% select(-c(sequencial_candidato, nome, estado_civil, cargo))

data2006_2010.correlation <- data2006_2010.correlation1  %>%
  mutate(uf = as.numeric(uf)) %>%
  mutate(partido = as.numeric(partido)) %>%
  mutate(sexo = as.numeric(sexo)) %>%
  mutate(grau = as.numeric(grau)) %>%
  mutate(ocupacao = as.numeric(ocupacao))
  

ggpairs(data2006_2010.correlation, axisLabels = "none")
```



```{r ,warning=FALSE, message=FALSE}

data2006_2010.lm = lm(data2006_2010$votos ~ data2006_2010$uf + data2006_2010$partido  + data2006_2010$quantidade_doacoes + data2006_2010$quantidade_doadores + data2006_2010$total_receita + data2006_2010$media_receita + data2006_2010$recursos_de_outros_candidatos.comites + data2006_2010$recursos_de_pessoas_fisicas + data2006_2010$recursos_de_pessoas_juridicas + data2006_2010$recursos_proprios + data2006_2010$recursos_de_partido_politico + data2006_2010$quantidade_despesas + data2006_2010$quantidade_fornecedores + data2006_2010$total_despesa + data2006_2010$media_despesa + data2006_2010$sexo + data2006_2010$grau + data2006_2010$estado_civil + data2006_2010$ocupacao + data2006_2010$ano)

summary(data2006_2010.lm)

```



#### 2006 and 2010
```{r}
data2006_2010.residuals = resid(data2006_2010.lm) 

densityplot(data2006_2010.residuals)
```

```{r ,warning=FALSE, message=FALSE}

data2006_2010.lm = lm(data2006_2010$votos ~  data2006_2010$quantidade_doacoes + data2006_2010$total_receita + data2006_2010$recursos_de_outros_candidatos.comites + data2006_2010$recursos_de_pessoas_fisicas + data2006_2010$recursos_de_pessoas_juridicas + data2006_2010$recursos_proprios+ data2006_2010$quantidade_despesas + data2006_2010$total_despesa + data2006_2010$media_despesa +  data2006_2010$ano)

summary(data2006_2010.lm)

```

