---
title: "Deputies Expenditures Analysis: Our Questions"
output: html_notebook
---

#TODO Here I gonna talk about the datas 
Link about it: http://www2.camara.leg.br/transparencia/acesso-a-informacao/copy_of_perguntas-frequentes/cota-para-o-exercicio-da-atividade-parlamentar 

For this lab we gonna need the following package to be installed
```{r}
install.packages("ggplot2")
install.packages("dplyr")
install.packages("tidyr")
install.packages("tidyverse")
```

Loading the libraries
```{r}
 library(ggplot2)
 library(dplyr)
 library(tidyr)
 library(tidyverse)

```

Set up the workspace folder

```{r}
setwd("~/git/data-analysis/Lab01")
```

Loading the data

```{r}
dados <- read_csv(("data/dadosCEAP.csv"))
```


```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
