---
title: "Election Prediction using knn, logistic regression, decision tree and adaboost as classifiers models"
author: "Jose Manoel Ferreira"
date: "December 03, 2018"
output:
  html_document:
    df_print: paged
---


#Introduction

<h3> Senario orveview: It is about to predict deputies election given a dataset whit some variables and the target is *situacao* which means if the deputie was elected (eleito) or not (nao_eleito). For that propouse we gonna: </h3>

>> Use four Classifier models

>> Train all models

>> Predict Results

>> Analyse Predictions 

>> Chosse the best model and predict for Kaggle Challenge https://www.kaggle.com/c/ufcg-cdp-20182-lab3/ 

Needed Libraries 
```{r ,warning=FALSE, message=FALSE}
library(caret)
library(mlbench)
library(C50)
library(dplyr)
library(plotly)
library(caret)
library(ROSE)
library(rpart)
library(GGally)
```


Setting up workspace

```{r}
setwd("~/git/data-analysis/lab03/")
```


Loading DATA

Our data frame will be the train.csv file, in which we'll peform predictions models
and test.csv will be used to Caggle challenge.

```{r}
data <- read.csv("data/all/train.csv")
test.kaggle <- read.csv("data/all/test.csv")
```

Here we gonna see the correlation between the variables, then will se the ones which has a strong correlation and remove, because keep both would be redundant for our prediction model.

```{r ,warning=FALSE, message=FALSE}
data.correlation1 <- data %>% select(-c(sequencial_candidato, nome, estado_civil, ano, cargo))

data.correlation <- data.correlation1  %>%
  mutate(situacao = as.numeric(situacao)) %>%
  mutate(uf = as.numeric(uf)) %>%
  mutate(partido = as.numeric(partido)) %>%
  mutate(sexo = as.numeric(sexo)) %>%
  mutate(grau = as.numeric(grau)) %>%
  mutate(ocupacao = as.numeric(ocupacao))

data.correlation %>% 
  select(-partido,
         -uf,-grau,-sexo) %>%
  na.omit() %>%
  ggcorr(palette = "RdBu",
         color = "grey50",
         label = TRUE, hjust = 1,
         label_size = 3, size = 4,
         nbreaks = 5, layout.exp = 7) +
  ggtitle("Correlation Between Variables")
```


We choosed to remove those three categoric variables in order to run the model, otherwise it would take too much time. But for a better result you could let them on the data. And also remove those variable which have strong correlation

```{r ,warning=FALSE, message=FALSE}
data <- data %>%
  select(-cargo, -nome, -ocupacao, -sexo, -total_despesa, -total_receita, -sequencial_candidato )
test.kaggle <- test.kaggle %>%
  select(-cargo, -nome, -ocupacao, total_despesa, -total_receita)
```

In the data would be better replace the NA for the column media, but we choosed replace by zero.


```{r}
data[is.na(data)] <- 0
test.kaggle[is.na(test.kaggle)] <- 0
```

As our target is to predict the variable *situacao* we need to see if our data is balanced, so what is the class distribution? 


```{r}
data_class_destribution <- data %>% group_by(situacao) %>% summarize(class_count = n())
p <- plot_ly(data_class_destribution, x = ~situacao, y = ~class_count, type = 'bar',
        marker = list(color = c('rgba(204,204,204,1)', 'rgba(222,45,38,0.8)'))) %>%
  layout(title = "Class Balance",
         xaxis = list(title = "Situation"),
         yaxis = list(title = "Count"))
p
```

<p>cleary unbalanced!</p>
<p>So what should we do? We gonna balance it.</p>
<p>There is some ways to balance data which are:</p>

 >> 1. Undersampling:
That method reduces the number of observation from the majoritary class in order to balance the data set.

>> 2. Oversampling:
This method increase the number of observation from the minoritary class and make it balanced.

>> 3. Both Sampling
Here it uses the technique 1 and 2 to make the data set balanced 

>> 4. ROSE Sampling:
Data synthetic generation and it provades a better stimation of original data.


Before balance it we gonna do a experiment. Let's create a model and see how is goes whitout balance in order to predict and see accuracy to compare in the future

For tu build our models we gonna need data to train and test so we'll divid the original data into train and test, 70% to raing and 30% to test.

```{r ,warning=FALSE, message=FALSE}
set.seed(42)
index <- createDataPartition(data$situacao, p = 0.7, list = FALSE)
unbalanced.train <- data[index, ]
unbalanced.test <- data[-index, ]
```


Decision Tree whit unbalanced data

```{r ,warning=FALSE, message=FALSE}

treeimb <- rpart(situacao ~ ., data = unbalanced.train)
pred.treeimb <- predict(treeimb, newdata = unbalanced.test)

metrics_unbalanced <- accuracy.meas(unbalanced.test$situacao, pred.treeimb[,2])
```

Surprisely we've got a good precision and recall. Anyways let's see how it goes whit balabced data.

Waht is test distribution?

```{r}
data_class_destribution <- unbalanced.test %>% group_by(situacao) %>% summarize(class_count = n())
p_test <- plot_ly(data_class_destribution, x = ~situacao, y = ~class_count, type = 'bar',
        marker = list(color = c('rgba(204,204,204,1)', 'rgba(222,45,38,0.8)'))) %>%
  layout(title = "Class Balance",
         xaxis = list(title = "Situation"),
         yaxis = list(title = "Count"))
p_test
```

Lets balance it, all data by using the 4th method ROSE Sampling which it gonna generate syntetich data.


```{r ,warning=FALSE, message=FALSE}
data.rose <- ROSE(situacao ~ ., data = data, seed = 1)$data
table(data.rose$situacao)
```
YEAH!

It looks pretty balanced now. That is great, so we now gonna peform some models and avaliate its metrics.




```{r}
new_index <- createDataPartition(data.rose$situacao, p = 0.7, list = FALSE)
new_train_data <- data.rose[index, ]
new_test_data  <- data.rose[-index, ]



```

```{r}
data_class_destribution <- new_test_data %>% group_by(situacao) %>% summarize(class_count = n())
p_test_bl <- plot_ly(data_class_destribution, x = ~situacao, y = ~class_count, type = 'bar',
        marker = list(color = c('rgba(204,204,204,1)', 'rgba(222,45,38,0.8)'))) %>%
  layout(title = "Class Balance",
         xaxis = list(title = "Situation"),
         yaxis = list(title = "Count"))
p_test_bl
```


Using the simple tree model to compare the results.

```{r}
new_treeimb <- rpart(situacao ~ ., data = new_train_data)
new_pred.treeimb <- predict(new_treeimb, newdata = new_test_data)


metrics_balanced <- accuracy.meas(new_test_data$situacao, new_pred.treeimb[,2])
```

So which are the precision for the tree model whith balanced data and unbalanced data?


Acuracioa: acertar
precisão: acertar sabendo

```{r}
dat <- data.frame(
    model = factor(c("Balanced","Unbalanced"), levels=c("Balanced","Unbalanced")),
    precision = c(metrics_balanced$precision, metrics_unbalanced$precision)
)

pl <- ggplot(data=dat, aes(x=model, y=precision, fill=model)) +
    geom_bar(colour="black", stat="identity") +
    guides(fill=FALSE)

pl <- ggplotly(pl)

pl
```
It seems that unbalanced is better, but it could be because the model trained got lucky, for that reason let's see how that models goes predicting data for balanced test.

```{r}
pred.treeimb_test <- predict(treeimb, newdata = new_test_data)
accuracy.meas(new_test_data$situacao, pred.treeimb_test[,2])
```
Unfortunately not that good. It was because the model had learned whit a unbalanced data which has a small observation as *eleito* and influenciated in the results.



Yes, we need to particionate our balanced data now, using the same schema before.

```{r}
set.seed(42)
index <- createDataPartition(data.rose$situacao, p = 0.7, list = FALSE)
train <- data.rose[index, ]
test <- data.rose[-index, ]
```

#knn
First model is Knn.

k-nearest neighbour classification for test set from training set. For each row of the test set, the k nearest (in Euclidean distance) training set vectors are found, and the classification is decided by majority vote, with ties broken at random. If there are ties for the kth nearest vector, all candidates are included in the vote.


```{r}
fitControl <- trainControl(method = "repeatedcv", 
                           number = 10,
                           repeats = 10)

preProcess = c("center", "scale","nzv" )
```

```{r}
model.knn <- train(situacao ~ ., 
               data = train,
               trControl = fitControl,
               method = "knn", # pode ser 'lasso'ldf
               metric = "Accuracy",
               preProcess = preProcess)

model.knn
```


```{r}
knn_prediction <- predict(model.knn,test)

knn_data <- data.frame(pred = knn_prediction, obs = test$situacao)

knn_cv <- round(defaultSummary(knn_data),digits = 4)

knn_cv
```

#Logistic Regression
Second model to be build.
That model aims to fit a regression curve, y= f(x), when y is a categorical variable.


```{r}
model.logistic_reg <- train(situacao ~ ., 
               data = train,
               trControl = fitControl,
               method = 'LogitBoost', 
               metric = "Accuracy",
               preProcess = preProcess)

model.logistic_reg
```



```{r}
logistic_reg_prediction <- predict(model.logistic_reg,test)

logistic_reg_data <- data.frame(pred = logistic_reg_prediction, obs = test$situacao)

logistic_reg_cv <- round(defaultSummary(logistic_reg_data),digits = 4)

logistic_reg_cv

```

  
#Decision Tree
Third model 
Decision tree is a graph to represent choices and their results in form of a tree. The nodes in the graph represent an event or choice and the edges of the graph represent the decision rules or conditions.

```{r}
new_index <- createDataPartition(data.rose$situacao, p = 0.7, list = FALSE)
new_train_data <- data.rose[index, ]
new_test_data  <- data.rose[-index, ]


new_treeimb <- rpart(situacao ~ ., data = new_train_data)
new_pred.treeimb <- predict(new_treeimb, newdata = new_test_data)


accuracy.meas(new_test_data$situacao, new_pred.treeimb[,2])
```

```{r}
model.tree_dec <- train(situacao ~ .,
                data= train, 
                method = "rpart",
                trControl = fitControl,
                cp=0.001,  
                metric = "Accuracy",
                maxdepth=20)
model.tree_dec
```

```{r}
tree_prediction <- predict(model.tree_dec,test)

tree_data <- data.frame(pred = tree_prediction, obs = test$situacao)

tree_cv <- round(defaultSummary(tree_data),digits = 4)

tree_cv

```


#AdaBoost
Boosting is an ensemble technique that attempts to create a strong classifier from a number of weak classifiers.

```{r}


model.adaboost <- train(situacao ~ ., 
               data = data.rose,
               trControl = trainControl(method = "repeatedcv", 
                           number = 10,
                           repeats = 5),
               method = 'adaboost', 
               metric = "Accuracy",
               preProcess = preProcess)

model.adaboost


```


```{r}
adaboost_prediction <- predict(model.adaboost,test)

adaboost_data <- data.frame(pred = adaboost_prediction, obs = test$situacao)

adaboost_cv <- round(defaultSummary(adaboost_data),digits = 4)

adaboost_cv

```

## Talking about metrics let's see Accuracy, Precision and F measure for each model


This method calculates the true positive, true negative, false positive and false negative. Then it will calculate the precision, recall and f measure.
```{r}
measurePrecisionRecall <- function(model_data){

  true_pos <- model_data %>%
    filter(obs == "eleito", pred == "eleito") %>% 
    nrow()
  true_neg <- model_data %>% 
    filter(obs == "nao_eleito" , pred == "nao_eleito" ) %>% 
    nrow()
  false_pos <- model_data %>% 
    filter(obs == "nao_eleito" , pred == "eleito") %>% 
    nrow()
  false_neg <- model_data %>% 
    filter(obs == "eleito", pred == "nao_eleito" ) %>% 
    nrow()
  
  precision <- true_pos / (true_pos + false_pos)
  recall <- true_pos / (true_pos + false_neg)
  fmeasure <- 2*(recall*precision)/(recall+precision)

  metrics <- list('precision' = precision, 'recall' = recall, 'fmeasure' = fmeasure)
  
  
  return(metrics)
}

```

```{r}
metrics.knn <- measurePrecisionRecall(knn_data)
metrics.logistic_reg <- measurePrecisionRecall(logistic_reg_data)
metrics.tree_dec <- measurePrecisionRecall(tree_data)
metrics.adaboost <- measurePrecisionRecall(adaboost_data)


```

Now, in the chart we see that the classifier Adaboos got excelent result, but it makes us questionate if there is no overfitting, probably. As well there is the metrics for other models, the second best was Logistic Regression followed by Tree Decision and then KNN.
```{r}
dat2 <- data.frame(
    model = factor(c("KNN", "KNN", "KNN", "Logic Regression", "Logic Regression", "Logic Regression", "Tree Decision", "Tree Decision", "Tree Decision", "Adabost", "Adabost", "Adabost")),
    metric = factor(c("Precision", "Recall", "F Measure", "Precision", "Recall", "F Measure", "Precision", "Recall", "F Measure", "Precision", "Recall", "F Measure")), levels=c("Precision", "Recall", "F Measure"),
    value = c(metrics.knn$precision, metrics.knn$recall, metrics.knn$fmeasure, metrics.logistic_reg$precision, metrics.logistic_reg$recall, metrics.logistic_reg$fmeasure, metrics.tree_dec$precision,  metrics.tree_dec$recall, metrics.tree_dec$fmeasure, metrics.adaboost$precision, metrics.adaboost$recall, metrics.adaboost$fmeasure)
)

p2 <- ggplot(data=dat2, aes(x=model, y=value, fill=metric)) +
    geom_bar(stat="identity", position=position_dodge(), colour="black") +
    scale_fill_manual(values=c("#999999", "#E69F00", "#7BCE94"))

p2 <- ggplotly(p2)

p2
```


##Which atribuite are most important to each model
As far we can see for *ano* and *sexo* we've a low outcome for importance, so those variables should be removed. As well the best variables pointe for all model are *recursos_de_pessoas_juridicas*, *recursos_de_pessoas_fisicas* and the other differ in the order.


###KNN
```{r}
varImp(model.knn)
```


###Logistic Regression

```{r}
varImp(model.logistic_reg)
```

###Decision Tree

```{r}
varImp(model.tree_dec)
```

###AdaBoost

```{r}
varImp(model.adaboost)
```


## Kaggle challenge
As far we can see for *ano* and *sexo* we've a low outcome for importance, so those variables should be removed.

As propose in the activite we are going to use our best model to submite the votos prediction to the challenge in Kaggle. 

```{r ,warning=FALSE, message=FALSE}
model.adaboost $xlevels[["ocupacao"]] <- union(model.adaboost$xlevels[["ocupacao"]], levels(test.kaggle$ocupacao))
prediction_ <- predict(model.adaboost , test.kaggle)
ID <- test.kaggle %>%
  select(sequencial_candidato)
colnames(ID)[colnames(ID)=="sequencial_candidato"] <- "ID"
predicted_file <- ID
predicted_file$Predicted <- prediction_
write.csv(predicted_file, "sample_submission.csv", row.names=FALSE)
```



usefull links:

http://www.treselle.com/blog/handle-class-imbalance-data-with-r/

https://www.analyticsvidhya.com/blog/2016/03/practical-guide-deal-imbalanced-classification-problems/

https://shiring.github.io/machine_learning/2017/04/02/unbalanced 



